Comparing vector-based and ACT-R memory models of tag retrieval:\protect\\User-customized hashtag/tag prediction on Twitter and StackOverflow

### 2014-07-28 writeups/dissertation/dissertation.tex
foo

\abstract{
  The growth of social media and user-created content on online sites provides unique opportunities to study models of declarative memory.
  The task of choosing a hashtag for a tweet and tagging a post on StackOverflow was framed as a declarative memory retrieval problem.
  Millions of posts and tweets were collected, and two declarative memory models were used to predict hashtags and tags that users choose on Twitter and StackOverflow.
  Two state-of-the-art cognitively-plausible memory models were evaluated on how accurate they predict a user's chosen tags: an ACT-R inspired Bayesian model and a random-permutation vector-based model.
  The results show that past user behavior of tag use is a strong predictor of future performance, and this past behavior was successfully incorporated into the random-permutation model that previously used only context.
  Also, the common natural-language processing method to attenuate low-predictor words in the dataset was linked to ACT-R's attentional weight term,
  and an entropy-weighting technique was developed that improved model accuracy by more efficiently allocating limited attentional resources to the predictive words in the posts and tweets.
  Word order was not found to be strong predictor of tag use, and the random-permutation model performed comparable to the Bayesian model without including word order,
  which shows that the strength of the random-permutation model is not in the ability to represent word order, but rather in the way in which context information is successfully compressed.
  The results of the large-scale exploration show how the architecture of the two memory models can be modified to significantly improve accuracy,
  and may suggest task-independent general modifications that can help improve model fit to human data in a much wider range of domains.
}

@@

The work for this proposal will expand on previous work that has tested retrieval models on large-scale datasets.
It will focus specifically on how two state-of-the-art declarative memory retrieval models (ACT-R Bayesian and random permutations) can be improved, by testing these models on two real-world tagging tasks.
One primary objective is to compare the random-permutation vector-based model to ACT-R's declarative memory theory when large datasets are used for each model. 
Another objective is to improve each model, and measure how incorporating the strengths of each model into the other (i.e., modifying model components) influences accuracy.
For example, it is currently unexplored how a vector-based model can be modified to retrieve results that are based not only on context, but also on the prior odds that a particular item in memory will be needed again.
Large-scale datasets will be used throughout the model development and evaluation process.
This provides an ideal testing environment for constraints and assumptions for each theory to be rapidly tested and new model components to be hypothesized and quickly evaluated.

@@

The core research question motivating this proposal is:
What kinds of cognitively-plausible models can best predict the human-created tags that a user on a social media site uses?
This question can be tested by generating a prediction for the most likely used hashtag whenever a user is about to generate a hashtag, and then comparing the user's chosen hashtag to the model's prediction.
This process can be framed as a memory retrieval problem.
That is, each user on a social media site has created content that contain a set of human-created tags within the context of each post.
The process of suggesting a relevant hashtag to a user can be thought of as a memory retrieval request for a hashtag, given prior hashtag use and current context.

@@

I will compare these two memory systems across two domains where users generate tags for posts: Twitter and StackOverflow.
Twitter is a microblogging service where users create 140 character tweets and broadcast those messages out to the users who follow their posts.
A user is free to annotate important concepts or words in the tweet by preceding words with hashtags ``\#''.
After the tweet is created the hashtags become hyperlinks, and if clicked on will take a user to a summary feed of all of the tweets across Twitter that have used that hashtag.
So these user-created hashtags on Twitter help connect the tweet to other tweets about the same topic through common hashtag use.
Some recent and often-used hashtags for Twitter are \emph{\#photography}, \emph{\#startup}, \emph{\#4change}, \emph{\#android}, and \emph{\#solar}.

@@

ACT-R's declarative memory and vector-based memory systems are substantially different in their mathematical structure.
Nonetheless, both can successfully model classic behavioral patterns found in word pairing experiments (e.g., \cite{Rutledge2008} modeling the ``fan effect'').
I will extend the work done by \citeauthor{Rutledge2008} by formally comparing these two models on large-scale hashtag prediction tasks.
I will also replace the vector-based model used by \citeauthor{Rutledge2008} with an improved and simplified random-permutation model \parencite{Sahlgren2008}.

An advantage of vector-based models is that word order can be easily incorporated into the single co-occurrence representation \parencite{Jones2007}.
Word-order information on sites like Twitter may contain highly-predictive pieces of information, as it is likely that specific words immediately precede specific hashtags.
However, it is unclear and unexplored how word order can be represented in the ACT-R declarative memory theory.
I will explore how the word-order strength of vector-based models can be incorporated into the ACT-R theory when testing the models on hashtag prediction tasks.

@@

The total activation ($A_{i}$) for a chunk in declarative memory is a function of two components: base level activation ($B_{i}$) and strength of association ($S_{ji}$).
The recall probability ($P_{i})$ that a chunk will be retrieved from memory increases with total activation ($A_{i}$).

@@

Base-level activation reflects the log prior odds of needing an observed chunk again.
The primary way to calculate these log prior odds is to use the standard base level equation in Table \ref{tabACTRModel}.
This equation formalizes how log prior odds are a function of both frequency and recency of prior exposure to a particular chunk.
Chunks used more frequently (either through exposure or from a retrieval) are more likely to be needed for retrieval again.
However, as time progresses and a particular chunk is no longer used, the activation for that chunk decays.
In this way the standard base level equation formalizes the time dynamics of the retrieval system, where a chunk's base-level activation evolves over time, depending on its current frequency and recency of use.

For some domains it is reasonable to assume that the base-level activations of each chunk within a particular time window of interest are stable, and do not change within that window.
Programming-language popularity over the past few years is a reasonable example of this.
Although the popularity of various programming languages has certainly changed slightly over the past few years (e.g., Clojure's growth), it is not the case that the changes have been drastic.
C\#, Python, and Java are still a few of the most popular languages, Common Lisp has not gained or lost much ground in popularity over the past few years.

Further, if this is a reasonable assumption to make, it greatly simplifies the computation of the base-level activation for chunks, and can turn the computation into a tractable problem for large datasets.
In these time-constant domains, one can compute the log prior odds of needing each chunk directly as the log odds ratio of exposure to that chunk compared to exposure to all other chunks.
For example, if the StackOverflow tag \emph{PHP} has been used four times as often as the tag \emph{Common Lisp}, then the prior odds of needing \emph{PHP} again is $\frac{.8}{.2}=4$ times that of \emph{Common Lisp}.

@@

Strength of association ($S_{ji}$) reflects the amount of log odds adjustment to the activation of a chunk, given the current context (i.e., external environment and internal state).
Context for the StackOverflow domain for example is represented as each word in the title and body of a post.
Context for the Twitter domain are the words in a tweet.
Association strength between a chunk in memory and a single contextual element can be computed directly by calculating its context-adjusted odds ratio:
The likelihood a chunk occurred with the current context ($p(i|j)$) over the likelihood that the chunk occurred in any of the other contexts ($p(i|\overline{j})$).
For large datasets, the likelihood that a chunk occurs in any particular context reaches near-zero values ($p(j) \Rightarrow 0$).
So it can be assumed that the likelihood that a chunk occurs in any context but one ($p(i|\overline{j})$) is equivalent to the likelihood that a chunk occurs in any context ($p(i)$).
This assumption has been both mentioned and used when deriving the $S_{ji}$ equation \parencite{Anderson1989}, as well as in recent research working with large-scale datasets \parencite{Stanley2013}.

@@

LSA uses the Singular Value Decomposition process to reduce the dimensionality of the word by document co-occurrence matrix.
The technique attempts to maximize the amount of variability in the data left after forcing the original dimensionality of the data to be reduced.
The best way to remove a dimension while maintaining the most variability is to look for words with highly-similar co-occurrence patterns and deduce that these words represent the same concept.
For example, synonyms such as \emph{\#photograph} and \emph{\#photo} would appear interchangeably across documents, so these terms would be collapsed to a single dimension and frequency counts would be pooled.
By pooling co-occurrence counts the model is able to generate better predictions with less data.

@@

When \textcite{Budiu2007} compared the simpler PMI (i.e., $S_{ji}$) measure to LSA on smaller TASA-corpus-size datasets, LSA performed much better (23\% vs. 60\%).
This was one of the primary motivating reasons that LSA was developed in the first place, since reduced dimensionality model out-performed the full dimensionality model on TASA-sized datasets.
However, once the dataset surpasses in size what is computationally feasible with LSA, the much simpler PMI measure can still be computed, and the performance for PMI becomes on par and even surpasses that of LSA.
\textcite{Budiu2007} also trained the PMI measure on the much larger Stanford WebBase corpus and tested the measure on the TOEFL.
This PMI model increased accuracy from 23\% to 51\% when training the dataset on the smaller TASA compared to the larger Stanford dataset.
Further, when \textcite{Turney2001} trained the PMI measure on the even larger AltaVista index of 350 million webpages, the accuracy on the TOEFL increased to 73.75\%, which is 10\% higher than the TASA-trained LSA.
This suggests that simpler models may be preferred to more complex (yet more accurate) models in large-scale domains, since the simpler models scale and can utilize all available information to generate predictions.

@@

One of the strengths of vector-based systems is that word order can naturally be represented alongside word co-occurrence in these distributed representations.
Representing word order is a case where both the encoding and decoding operations are applied.
As a motivating example, say that one wants to encode the first and second words immediately preceding a hashtag in a tweet.
Simply counting co-occurrences of the two immediately-preceding words and hashtags does not include the order information of the words in the representation.
With this representation, it would be impossible to tell the difference between 1-back words that immediately preceded the hashtag, and 2-back words that occurred just prior to these 1-back words.
Order information is lost if a simply co-occurrence counting technique is used.

@@

\textcite{Jones2007} created the BEAGLE vector-based memory representation.
Each memory vector ($m_{i}$) is a summed representation of unordered and ordered word co-occurrences ($c_{i} + o_{i}$).
This model uses circular convolution and circular correlation as a way to encode and decode word order.
BEAGLE performed similar to LSA when trained on the TASA and tested on the TOEFL (55.6\% compared to 55.3\%).
Also, BEAGLE with unordered context and word order was more correlated to the WordNet synonym database than LSA (-.311 compared to -.165).
Further, it was shown that incorporating word-order information into the same representation resulted in minimal data loss (less than 1 percent of total predictive variance).
This model provided an efficient algorithm for using a single representation for unordered and ordered information that showed performance similar to if not better than LSA.

@@

\textcite{Recchia2010} found that random permutations could use smaller vectors than circular convolution approaches (e.g., BEAGLE) to produce equivalent performance (better compression),
and also scale to larger datasets (better computational efficiency).
The vector length by performance tradeoff was tested on a simple paired associative retrieval task. 
Retrieval accuracy for the random-permutation vector-based models remained high relative to the circular convolution model even after reducing each environment vector's length from \num{2048} to \num{1024}.
This indicates that vector-based representations using random permutations compress better than those using circular convolution.

To test computational efficiency, each model was trained on a subset of the Wikipedia corpus, and tested on the Test of English as a Foreign Language (TOEFL) assessment (among other tasks).
The random-permutation model was trained on \num{2.33} GB corpus.
The BEAGLE model was unable to train on a corpus this large, so both models were also trained on an additional \num{35} MB subset of the Wikipedia corpus.
The random permutation model performed as well or slightly better than the BEAGLE model when trained on the smaller corpus.
The model's performance also significantly improved when trained on the much larger corpus that was computationally intractable for the BEAGLE model.
These results make a strong case to favor random-permutations over circular convolution when implementing vector-based memory representations.

@@

This is a counterintuitive result.
It seems plausible that an SVD technique that takes into account word similarity when reducing the matrix should generate a reduced matrix that has more information than a randomly-compressed matrix.
This does not seem to be the case in testing however, as vector-based approaches like BEAGLE and random permutations perform just as well as the LSA when trained and tested on similar datasets.
Nonetheless, it has been suggested to perform an SVD reduction technique on a compressed random-permutation vector \parencite{Kanerva2000}.
This approach might be worth investigating, as it would reduce computational requirements even further, and possibly improve performance by pooling co-occurrence counts from latent topics together.

@@

One strength of a vector-based model over ACT-R's Bayesian DM model is that word order can be naturally represented alongside unordered co-occurrence information.
With the ACT-R model, for example, it is unclear how to represent that a particular word appeared just before a hashtag. 
It may be that multiple co-occurrence matrices must be used to represent word-order co-occurrences and unordered co-occurrences separately.
However, it would be more parsimonious if these matrices could be combined into a single matrix representation, much like how this information is combined for vector-based models.

@@

The Netflix Prize \parencite{Bennett2007} is an example of a recommendation system where prior user behavior and current context was used to present the most relevant information to the user.
Netflix is an online-subscription service where users can stream their favorite television episodes and movies.
Each user of the site is only interested in a small portion of the entire library of content that Netflix provides.
In order to provide a high-quality user experience on the site, Netflix created a content-recommendation service.
This tool recommends to the user movies and television episodes that most likely match their specific interests.

Netflix wanted to improve the accuracy of their recommendation system, so they held a competition where teams could try to improve on the accuracy of the service.
The winning team utilized a statistical model that performed a type of factor analysis to identify important dimensions in the data (e.g., drama, action, cult classic) and place values for each user along each dimension. 
This method used prior user behavior to classify a user along the dimensions of the reduced space, and current user behavior (e.g., just watched) to understand what they were interested in watching at that moment.
By using a factor-analysis technique and combining these indicators, the model improved on Netflix's original recommendation system by 10\%.

@@

Lucene \parencite{McCandless2010} is a commonly-used open-source software tool that measures the similarity between a query and candidate documents in a database.
It uses the term frequency---inverse document frequency co-occurrence measure to quantify the similarity between a word in the query and a possible document to retrieve.

@@

The activation of each tag ($A_{i}$) is a function of four components:
That tag's base level activation ($B_{i}$), contextual activation for the words in the title, activation for words in the body, and an offset term ($O$).
This model has been modified from the standard ACT-R model (Table \ref{tabACTRModel}) in three ways:
Two separate contextual components were used (words in title and body of post), an offset term was added, and an entropy measure was used to compute the attentional weight ($W_{j}$).

@@

Also, a vector-based retrieval model has not yet been tested on Twitter.
One of the advantages of vector-based models is that word order can be easily incorporated into the single co-occurrence representation \parencite{Jones2007}.
It seems quite likely that words used just before a hashtag in a tweet (combined with their position) will be highly predictive of the hashtag chosen.
To our knowledge, word order has not yet been taken into account for any of the Twitter hashtag-prediction models that have already been tested.
It will be interesting to see how much (if at all) accuracy improves when incorporating word-order information into the model.

@@

As some preliminary research, I have developed and tested a user-customized model for StackOverflow by modifying the model in \textcite{Stanley2013}.
To incorporate the user's prior tag history into the model, the base-level activation term was modified from the original equation in ACT-R.
$B_{i}$ is based on log odds occurrence, which in turn is based on frequency of tag use, or probability of tag use $p_{i}$.
To incorporate a user's prior, a weighted average for $p_{i}$ was used,
where the probability of using tag $i$ ranges from the global probability ($pGlobal_{i}$) to the user's specific probability ($pUser_{i}$) as the user's total tag count increases.
Overall model accuracy improved by 5\% after incorporating the user's prior tag history into the model.

This preliminary research is encouraging, and suggests that model fit can improve when tag predictions are customized to the specific user's past tagging history.
It also seems plausible that model accuracy can improve an even greater amount by
[1] figuring out the optimal way to combine global tag history with a user's past tagging history, and
[2] using the time-dependent form of the base level component in Table \ref{tabACTRModel}, in order to account for the fact that a user's tagging preferences change over time.
So these proposed modifications will be explored further when incorporating a user's past tagging history into the models for this research.

@@

\textcite{Owoputi2013}'s algorithm handles basic word stemming (e.g., removing 'ed' and 'ing' endings), but
with the StackOverflow dataset, it is also possible to convert synonym tags to canonical base tags, since the community maintains a tag-synonym database.
This can be thought of as semantic stemming, where two words with identical meaning are converted to an agreed-upon canonical word.
So tag words in the post and used tags for the post were converted to their root tag when possible.

@@

Stop words are commonly removed by either filtering based on a pre-determined list or filtering based on the high-frequency words that occur in the dataset.
\textcite{Sahlgren2008} tried both methods to remove stop words for the random-permutation vector-based model, and found that the data-driven approach
(i.e., building the list based on the high-frequency words that occurred in the dataset) worked much better than using a pre-determined list.

@@

\begin{table}[!ht]
  \caption{Base-level component of ACT-R declarative memory}
  \label{tabACTRBLLModel}
  {\tabulinesep=1.2mm
    \begin{tabu}{ll}
      \hline
      Common Name &  Equation \\
      \hline
      Base Level Activation &	$B_{i} = log \sum_{j=1}^{n} {t_{j}}^{-d}$ \\
      Optimized Learning &	$B_{i} = log \frac{n}{1-d} - d * log L$ \\
      \hline
    \end{tabu}
  }
\end{table}

@@

For both equations, $i$ represents each chunk (e.g., a word), $j$ is a specific observation for a chunk (e.g., the 2nd time that ``a'' appears), and $n$ is the number of times each chunk has been observed.
Since the optimized learning form assumes a constant presentation rate for each chunk,
it only has to keep track of the time since the first presentation of a chunk $L$ and the total number of instances observed for that chunk. 
It does not have to keep track of the time each specific chunk instance was observed.
Consequently, the optimized learning form of $B_{i}$ is computationally much less expensive than the standard form, and is actually set to the default in the ACT-R common-lisp implementation.

@@

The base-level ($B_{i}$) component in Table \ref{tabACTRBLLModel} is different than the $B_{i}$ used in \textcite{Stanley2013} (Table \ref{tabModACTRModel}).
\citeauthor{Stanley2013} assumed no time dependence, no decay rate, and that the $B_{i}$ is simply a log odds computation of the probability that each chunk is likely to be presented again, given past frequency.
If a decay rate of 0 is used on the optimized learning form of $B_{i}$, then the equation collapses to $log n_{i}$,
which is a nonlinear transformation of the $B_{i}$ used by \citeauthor{Stanley2013} ($log \frac{p_{i}}{1-p_{i}}$).
Both forms maintain the same rank ordering, so they should produce similar results.
So the $B_{i}$ component used by \citeauthor{Stanley2013} can be roughly thought of as the standard form of $B_{i}$ when the decay rate is 0 (i.e., a pure frequency model).

@@

Model accuracy relaxed across posts with and without the offset model component was evaluated for both the Bayesian and random-permutation model.

@@

Model accuracy results after incorporating word order into the random-permutation model are included in Figure \ref{figContextOrder}

>>>>

\abstract{
  The growth of social media and user-created content on online sites provides unique opportunities to study models of declarative memory.
  The task of choosing a hashtag for a tweet and tagging a post on StackOverflow was framed as a declarative memory retrieval problem.
  Millions of posts and tweets were collected, and two declarative memory models were used to predict hashtags and tags that users choose on Twitter and StackOverflow.
  Two state-of-the-art cognitively-plausible memory models were evaluated on how accurate they predict a user's chosen tags: an ACT-R inspired Bayesian model and a random permutation vector-based model.
  The results show that past user behavior of tag use is a strong predictor of future performance, and this past behavior was successfully incorporated into the random permutation model that previously used only context.
  Also, the common natural-language processing method to attenuate low-predictor words in the dataset was linked to ACT-R's attentional weight term,
  and an entropy-weighting technique was developed that improved model accuracy by more efficiently allocating limited attentional resources to the predictive words in the posts and tweets.
  Word order was not found to be strong predictor of tag use, and the random permutation model performed comparable to the Bayesian model without including word order,
  which shows that the strength of the random permutation model is not in the ability to represent word order, but rather in the way in which context information is successfully compressed.
  The results of the large-scale exploration show how the architecture of the two memory models can be modified to significantly improve accuracy,
  and may suggest task-independent general modifications that can help improve model fit to human data in a much wider range of domains.
}

@@

The work for this proposal will expand on previous work that has tested retrieval models on large-scale datasets.
It will focus specifically on how two state-of-the-art declarative memory retrieval models (ACT-R Bayesian and random permutations) can be improved, by testing these models on two real-world tagging tasks.
One primary objective is to compare the random permutation vector-based model to ACT-R's declarative memory theory when large datasets are used for each model. 
Another objective is to improve each model, and measure how incorporating the strengths of each model into the other (i.e., modifying model components) influences accuracy.
For example, it is currently unexplored how a vector-based model can be modified to retrieve results that are based not only on context, but also on the prior odds that a particular item in memory will be needed again.
Large-scale datasets will be used throughout the model development and evaluation process.
This provides an ideal testing environment for constraints and assumptions for each theory to be rapidly tested and new model components to be hypothesized and quickly evaluated.

@@

The core research question motivating this proposal is:
What kinds of cognitively-plausible models can best predict the tags that a user on a social media site uses?
This question can be tested by generating a prediction for the most likely used hashtag whenever a user is about to generate a hashtag, and then comparing the user's chosen hashtag to the model's prediction.
This process can be framed as a memory retrieval problem.
That is, each user on a social media site has created content that contain a set of tags within the context of each post.
The process of suggesting a relevant hashtag to a user can be thought of as a memory retrieval request for a hashtag, given prior hashtag use and current context.

@@

I will compare these two memory systems across two domains where users generate tags for posts: Twitter and StackOverflow.
Twitter is a microblogging service where users create 140 character tweets and broadcast those messages out to the users who follow their posts.
A user is free to annotate important concepts or words in the tweet by preceding words with hashtags ``\#''.
After the tweet is created the hashtags become hyperlinks, and if clicked on will take a user to a summary feed of all of the tweets across Twitter that have used that hashtag.
So these user-created hashtags on Twitter help connect the tweet to other tweets about the same topic through common hashtag use.
Some recent and often used hashtags for Twitter are \emph{\#photography}, \emph{\#startup}, \emph{\#4change}, \emph{\#android}, and \emph{\#solar}.

@@

ACT-R's declarative memory and vector-based memory systems are substantially different in their mathematical structure.
Nonetheless, both can successfully model classic behavioral patterns found in word pairing experiments (e.g., \cite{Rutledge2008} modeling the ``fan effect'').
I will extend the work done by \citeauthor{Rutledge2008} by formally comparing these two models on large-scale hashtag prediction tasks.
I will also replace the vector-based model used by \citeauthor{Rutledge2008} with an improved and simplified random permutation model \parencite{Sahlgren2008}.

An advantage of vector-based models is that word order can be easily incorporated into the single co-occurrence representation \parencite{Jones2007}.
Word order information on sites like Twitter may contain highly predictive pieces of information, as it is likely that specific words immediately precede specific hashtags.
However, it is unclear and unexplored how word order can be represented in the ACT-R declarative memory theory.
I will explore how the word-order strength of vector-based models can be incorporated into the ACT-R theory when testing the models on hashtag prediction tasks.

@@

The total activation ($A_{i}$) for a chunk in declarative memory is a function of two components: base-level activation ($B_{i}$) and strength of association ($S_{ji}$).
The recall probability ($P_{i})$ that a chunk will be retrieved from memory increases with total activation ($A_{i}$).

@@

Base-level activation reflects the log prior odds of needing an observed chunk again.
The primary way to calculate these log prior odds is to use the standard base-level equation in Table \ref{tabACTRModel}.
This equation formalizes how log prior odds are a function of both frequency and recency of prior exposure to a particular chunk.
Chunks used more frequently (either through exposure or from a retrieval) are more likely to be needed for retrieval again.
However, as time progresses and a particular chunk is no longer used, the activation for that chunk decays.
In this way the standard base-level equation formalizes the time dynamics of the retrieval system, where a chunk's base-level activation evolves over time, depending on its current frequency and recency of use.

For some domains it is reasonable to assume that the base-level activations of each chunk within a particular time window of interest are stable, and do not change within that window.
Programming language popularity over the past few years is a reasonable example of this.
Although the popularity of various programming languages has certainly changed slightly over the past few years (e.g., Clojure's growth), it is not the case that the changes have been drastic.
C\#, Python, and Java are still a few of the most popular languages, Common Lisp has not gained or lost much ground in popularity over the past few years.

Further, if this is a reasonable assumption to make, it greatly simplifies the computation of the base-level activation for chunks, and can turn the computation into a tractable problem for large datasets.
In these time-constant domains, one can compute the log prior odds of needing each chunk directly as the log-odds ratio of exposure to that chunk compared to exposure to all other chunks.
For example, if the StackOverflow tag \emph{PHP} has been used four times as often as the tag \emph{Common Lisp}, then the prior odds of needing \emph{PHP} again is $\frac{.8}{.2}=4$ times that of \emph{Common Lisp}.

@@

Strength of association ($S_{ji}$) reflects the amount of log-odds adjustment to the activation of a chunk, given the current context (i.e., external environment and internal state).
Context for the StackOverflow domain for example is represented as each word in the title and body of a post.
Context for the Twitter domain are the words in a tweet.
Association strength between a chunk in memory and a single contextual element can be computed directly by calculating its context-adjusted odds ratio:
The likelihood a chunk occurred with the current context ($p(i|j)$) over the likelihood that the chunk occurred in any of the other contexts ($p(i|\overline{j})$).
For large datasets, the likelihood that a chunk occurs in any particular context reaches near-zero values ($p(j) \Rightarrow 0$).
So it can be assumed that the likelihood that a chunk occurs in any context but one ($p(i|\overline{j})$) is equivalent to the likelihood that a chunk occurs in any context ($p(i)$).
This assumption has been both mentioned and used when deriving the $S_{ji}$ equation \parencite{Anderson1989}, as well as in recent research working with large-scale datasets \parencite{Stanley2013}.

@@

LSA uses the Singular Value Decomposition process to reduce the dimensionality of the word by document co-occurrence matrix.
The technique attempts to maximize the amount of variability in the data left after forcing the original dimensionality of the data to be reduced.
The best way to remove a dimension while maintaining the most variability is to look for words with highly similar co-occurrence patterns and deduce that these words represent the same concept.
For example, synonyms such as \emph{\#photograph} and \emph{\#photo} would appear interchangeably across documents, so these terms would be collapsed to a single dimension and frequency counts would be pooled.
By pooling co-occurrence counts the model is able to generate better predictions with less data.

@@

When \textcite{Budiu2007} compared the simpler PMI (i.e., $S_{ji}$) measure to LSA on smaller TASA-sized datasets, LSA performed much better (23\% vs. 60\%).
This was one of the primary motivating reasons that LSA was developed in the first place, since reduced dimensionality model outperformed the full dimensionality model on TASA-sized datasets.
However, once the dataset surpasses in size what is computationally feasible with LSA, the much simpler PMI measure can still be computed, and the performance for PMI becomes on par and even surpasses that of LSA.
\textcite{Budiu2007} also trained the PMI measure on the much larger Stanford WebBase corpus and tested the measure on the TOEFL.
This PMI model increased accuracy from 23\% to 51\% when training the dataset on the smaller TASA compared to the larger Stanford dataset.
Further, when \textcite{Turney2001} trained the PMI measure on the even larger AltaVista index of 350 million webpages, the accuracy on the TOEFL increased to 73.75\%, which is 10\% higher than the TASA-trained LSA.
This suggests that simpler models may be preferred to more complex (yet more accurate) models in large-scale domains, since the simpler models scale and can utilize all available information to generate predictions.

@@

One of the strengths of vector-based systems is that word order can naturally be represented alongside word co-occurrence in these distributed representations.
Representing word order is a case where both the encoding and decoding operations are applied.
As a motivating example, say that one wants to encode the first and second words immediately preceding a hashtag in a tweet.
Simply counting co-occurrences of the two immediately preceding words and hashtags does not include the order information of the words in the representation.
With this representation, it would be impossible to tell the difference between 1-back words that immediately preceded the hashtag, and 2-back words that occurred just prior to these 1-back words.
Order information is lost if a simply co-occurrence counting technique is used.

@@

\textcite{Jones2007} created the BEAGLE vector-based memory representation.
Each memory vector ($m_{i}$) is a summed representation of unordered and ordered word co-occurrences ($c_{i} + o_{i}$).
This model uses circular convolution and circular correlation as a way to encode and decode word order.
BEAGLE performed similar to LSA when trained on the TASA and tested on the TOEFL (55.6\% compared to 55.3\%).
Also, BEAGLE with unordered context and word order was more correlated to the WordNet synonym database than LSA (-.311 compared to -.165).
Further, it was shown that incorporating word order information into the same representation resulted in minimal data loss (less than 1 percent of total predictive variance).
This model provided an efficient algorithm for using a single representation for unordered and ordered information that showed performance similar to if not better than LSA.

@@

\textcite{Recchia2010} found that random permutations could use smaller vectors than circular convolution approaches (e.g., BEAGLE) to produce equivalent performance (better compression),
and also scale to larger datasets (better computational efficiency).
The vector length by performance tradeoff was tested on a simple paired associative retrieval task. 
Retrieval accuracy for the random permutation vector-based models remained high relative to the circular convolution model even after reducing each environment vector's length from \num{2048} to \num{1024}.
This indicates that vector-based representations using random permutations compress better than those using circular convolution.

To test computational efficiency, each model was trained on a subset of the Wikipedia corpus, and tested on the Test of English as a Foreign Language (TOEFL) assessment (among other tasks).
The random permutation model was trained on \num{2.33} GB corpus.
The BEAGLE model was unable to train on a corpus this large, so both models were also trained on an additional \num{35} MB subset of the Wikipedia corpus.
The random permutation model performed as well or slightly better than the BEAGLE model when trained on the smaller corpus.
The model's performance also significantly improved when trained on the much larger corpus that was computationally intractable for the BEAGLE model.
These results make a strong case to favor random permutations over circular convolution when implementing vector-based memory representations.

@@

This is a counterintuitive result.
It seems plausible that an SVD technique that takes into account word similarity when reducing the matrix should generate a reduced matrix that has more information than a randomly-compressed matrix.
This does not seem to be the case in testing however, as vector-based approaches like BEAGLE and random permutations perform just as well as the LSA when trained and tested on similar datasets.
Nonetheless, it has been suggested to perform an SVD reduction technique on a compressed random permutation vector \parencite{Kanerva2000}.
This approach might be worth investigating, as it would reduce computational requirements even further, and possibly improve performance by pooling co-occurrence counts from latent topics together.

@@

One strength of a vector-based model over ACT-R's Bayesian DM model is that word order can be naturally represented alongside unordered co-occurrence information.
With the ACT-R model, for example, it is unclear how to represent that a particular word appeared just before a hashtag. 
It may be that multiple co-occurrence matrices must be used to represent ordered and unordered co-occurrences separately.
However, it would be more parsimonious if these matrices could be combined into a single matrix representation, much like how this information is combined for vector-based models.

@@

The Netflix Prize \parencite{Bennett2007} is an example of a recommendation system where prior user behavior and current context was used to present the most relevant information to the user.
Netflix is an online subscription service where users can stream their favorite television episodes and movies.
Each user of the site is only interested in a small portion of the entire library of content that Netflix provides.
In order to provide a high quality user experience on the site, Netflix created a content-recommendation service.
This tool recommends to the user movies and television episodes that most likely match their specific interests.

Netflix wanted to improve the accuracy of their recommendation system, so they held a competition where teams could try to improve on the accuracy of the service.
The winning team utilized a statistical model that performed a type of factor analysis to identify important dimensions in the data (e.g., drama, action, cult classic) and place values for each user along each dimension. 
This method used prior user behavior to classify a user along the dimensions of the reduced space, and current user behavior (e.g., just watched) to understand what they were interested in watching at that moment.
By using a factor analysis technique and combining these indicators, the model improved on Netflix's original recommendation system by 10\%.

@@

Lucene \parencite{McCandless2010} is a commonly-used open source software tool that measures the similarity between a query and candidate documents in a database.
It uses the term frequency---inverse document frequency co-occurrence measure to quantify the similarity between a word in the query and a possible document to retrieve.

@@

The activation of each tag ($A_{i}$) is a function of four components:
That tag's base-level activation ($B_{i}$), contextual activation for the words in the title, activation for words in the body, and an offset term ($O$).
This model has been modified from the standard ACT-R model (Table \ref{tabACTRModel}) in three ways:
Two separate contextual components were used (words in title and body of post), an offset term was added, and an entropy measure was used to compute the attentional weight ($W_{j}$).

@@

Also, a vector-based retrieval model has not yet been tested on Twitter.
One of the advantages of vector-based models is that word order can be easily incorporated into the single co-occurrence representation \parencite{Jones2007}.
It seems quite likely that words used just before a hashtag in a tweet (combined with their position) will be highly predictive of the hashtag chosen.
To our knowledge, word order has not yet been taken into account for any of the Twitter hashtag-prediction models that have already been tested.
It will be interesting to see how much (if at all) accuracy improves when incorporating word order information into the model.

@@

As some preliminary research, I have developed and tested a user-customized model for StackOverflow by modifying the model in \textcite{Stanley2013}.
To incorporate the user's prior tag history into the model, the base-level activation term was modified from the original equation in ACT-R.
$B_{i}$ is based on log-odds occurrence, which in turn is based on frequency of tag use, or probability of tag use $p_{i}$.
To incorporate a user's prior, a weighted average for $p_{i}$ was used,
where the probability of using tag $i$ ranges from the global probability ($pGlobal_{i}$) to the user's specific probability ($pUser_{i}$) as the user's total tag count increases.
Overall model accuracy improved by 5\% after incorporating the user's prior tag history into the model.

This preliminary research is encouraging, and suggests that model fit can improve when tag predictions are customized to the specific user's past tagging history.
It also seems plausible that model accuracy can improve an even greater amount by
[1] figuring out the optimal way to combine global tag history with a user's past tagging history, and
[2] using the time-dependent form of the base-level component in Table \ref{tabACTRModel}, in order to account for the fact that a user's tagging preferences change over time.
So these proposed modifications will be explored further when incorporating a user's past tagging history into the models for this research.

@@

\textcite{Owoputi2013}'s algorithm handles basic word stemming (e.g., removing 'ed' and 'ing' endings), but
with the StackOverflow dataset, it is also possible to convert synonym tags to canonical base tags, since the community maintains a tag-synonym database.
This can be thought of as semantic stemming, where two words with identical meaning are converted to the same canonical word.
So tag words in the post and used tags for the post were converted to their root tag when possible.

@@

Stop words are commonly removed by either filtering based on a pre-determined list or filtering based on the high-frequency words that occur in the dataset.
\textcite{Sahlgren2008} tried both methods to remove stop words for the random permutation vector-based model, and found that the data-driven approach
(i.e., building the list based on the high-frequency words that occurred in the dataset) worked much better than using a pre-determined list.

@@

\begin{table}[!ht]
  \caption{Base-level component of ACT-R declarative memory}
  \label{tabACTRBLLModel}
  {\tabulinesep=1.2mm
    \begin{tabu}{ll}
      \hline
      Common Name &  Equation \\
      \hline
      Base-Level Activation &	$B_{i} = log \sum_{j=1}^{n} {t_{j}}^{-d}$ \\
      Optimized Learning &	$B_{i} = log \frac{n}{1-d} - d * log L$ \\
      \hline
    \end{tabu}
  }
\end{table}

@@

For both equations, $i$ represents each chunk (e.g., a word), $j$ is a specific observation for a chunk (e.g., the 2nd time that ``a'' appears), and $n$ is the number of times each chunk has been observed.
Since the optimized learning form assumes a constant presentation rate for each chunk,
it only has to keep track of the time since the first presentation of a chunk $L$ and the total number of instances observed for that chunk. 
It does not have to keep track of the time each specific chunk instance was observed.
Consequently, the optimized learning form of $B_{i}$ is computationally much less expensive than the standard form, and is actually set to the default in the ACT-R Common Lisp implementation.

@@

The base-level ($B_{i}$) component in Table \ref{tabACTRBLLModel} is different than the $B_{i}$ used in \textcite{Stanley2013} (Table \ref{tabModACTRModel}).
\citeauthor{Stanley2013} assumed no time dependence, no decay rate, and that the $B_{i}$ is simply a log-odds computation of the probability that each chunk is likely to be presented again, given past frequency.
If a decay rate of 0 is used on the optimized learning form of $B_{i}$, then the equation collapses to $log n_{i}$,
which is a nonlinear transformation of the $B_{i}$ used by \citeauthor{Stanley2013} ($log \frac{p_{i}}{1-p_{i}}$).
Both forms maintain the same rank ordering, so they should produce similar results.
So the $B_{i}$ component used by \citeauthor{Stanley2013} can be roughly thought of as the standard form of $B_{i}$ when the decay rate is 0 (i.e., a pure frequency model).

@@

Model accuracy relaxed across posts with and without the offset model component was evaluated for both the Bayesian and random permutation model.

@@

Model accuracy results after incorporating word order into the random permutation model are included in Figure \ref{figContextOrder}
